A Constraint Solving Approach to Parikh Images of Regular Languages

### 2023-06-27 multiple.tex
Progress on example

We reuse the transition variable associations of our previous
\cref{sec:single:example}, from \cref{eq:single-filter}, but add the definitions
for $B$ shown in \cref{eq:multi-filter}:
\begin{equation}\label{eq:multi-filter}
    \begin{aligned}
        \Filter(\FromLabelTo{A}{a / l_{B+}}{F_B} ) = \TransitionVar_6\\
        \Filter(\FromLabelTo{A}{c / l_{B+}}{A} ) = \TransitionVar_7\\
        \Filter(\FromLabelTo{B}{b / l_{B+}}{S})= \TransitionVar_8 \\
        \Filter(\FromLabelTo{B}{c / l_{B+}}{B})= \TransitionVar_9\\
        \Filter(\FromLabelTo{B}{c / l_{B+}}{F_B})= \TransitionVar_{10}\\
        \Filter(\FromLabelTo{I_B}{a / l_{B+}}{A})= \TransitionVar_{11}\\
        \Filter(\FromLabelTo{I_B}{b / l_{B+}}{B})= \TransitionVar_{12}\\
        \FlowEq(B, \Filter) =  % FIXME FLOWS FOR B \\
    \end{aligned}
\end{equation}

@@

\begin{figure}
  \centering
\begin{prooftree}
  \hypo{
    {\begin{array}{lll}
      1 = \TransitionVar_{10} & \land \TransitionVar_3 = \TransitionVar_{11} & \land 1 = \TransitionVar{12} \\ 
      \land 1 = \TransitionVar_{11} &  & \land 2 > 0
    \end{array}}  
  }
  \infer1[\Subsume{}, \Expand{}, \Subsume{}]{
    {\begin{array}{ll}
      2 > 0 \land 1 = \TransitionVar_{10} & \land \TransitionVar_3 = \TransitionVar_{11} \\
      \land 1 = \TransitionVar_{12} & \land 1 = \TransitionVar_{11} \\
      \land \Image{}_{\Automaton', \Map}(\Filter, \VectorLiteral{2, 0, 1}) & \land \Connected(\SomethingCSomething{}, \Filter) \\
    \end{array}}  
  }
  \infer1[\EquationReasoning{}]{
  {  \begin{array}{ll}
      1 = \TransitionVar_{10} \land \TransitionVar_3 = \TransitionVar_{11} & 
      \land 1 = \TransitionVar_{12} \land 1 = \TransitionVar_{11} \\
      \land \Image{}_{\Automaton', \Map}(\Filter, 
      \VectorLiteral{
        \TransitionVar_{7a} + \TransitionVar_{9a},
        0,
        \TransitionVar_{7c} + \TransitionVar_{9c} + 1}) & \land 
        \TransitionVar_6 + \TransitionVar_4 = 0 \\
        \land 1 - \TransitionVar_6 > 0 \land \TransitionVar_6 = 0 &
        \land \TransitionVar_4 = 0 \land  \TransitionVar_3 > 0 \\
        \land 2\TransitionVar_6 + \TransitionVar_4 = 0 & \land 
          \VectorLiteral{
        \TransitionVar_{7a} + \TransitionVar_{9a},
        0,
        \TransitionVar_{7c} + \TransitionVar_{9c} + 1} 
        = \VectorLiteral{
          2,
          0,
          \TransitionVar_3} \\ \land 
      \Connected(\SomethingCSomething{}, \Filter) &
      \land \TransitionVar_{7a} + \TransitionVar_{9a} > 0
    \end{array} } 
  }
  \infer1[\Materialise]{
    {\begin{array}{ll}
      \TransitionVar_6 + \TransitionVar_4 = 0 \land 1 - \TransitionVar_6 > 0 &
      \land \TransitionVar_6 = 0  \land \TransitionVar_4 = 0 \\
      \land \TransitionVar_3 > 0 & \land 2\TransitionVar_6 + \TransitionVar_4 = 0 \\
      \land 
      \VectorLiteral{
        \TransitionVar_{7a} + \TransitionVar_{9a},
        0,
        \TransitionVar_{7c} + \TransitionVar_{9c} + 1}
        = \VectorLiteral{
          2,
          0,
          \TransitionVar_3} & \land
      \Connected(\SomethingCSomething{}, \Filter) \\
      \land 
      \Image{}_{\Tuple{\AcaOrBc{},\SomethingCSomething{}}, \Map}(\Filter, 
      \VectorLiteral{\TransitionVar_{7a} + \TransitionVar_{9a}, 0, \TransitionVar_{7c} + \TransitionVar_{9c} + 1}) & \land \TransitionVar_{7a} + \TransitionVar_{9a} > 0
    \end{array}}
  }
  \infer1[\Subsume{}]{
  {\begin{array}{ll}
    \TransitionVar_6 + \TransitionVar_4 = 0 \land 
    1 - \TransitionVar_6 > 0 &
    \land
    \TransitionVar_6 = 0 \land 
    \TransitionVar_4 = 0 \\
    \land
    \TransitionVar_3 > 0 & \land
    2\TransitionVar_6 + \TransitionVar_4 = 0 \\ 
    \land
    \VectorLiteral{
      \TransitionVar_{7a} + \TransitionVar_{9a},
      0,
      \TransitionVar_{7c} + \TransitionVar_{9c} + 1}
      = \VectorLiteral{
        2,
        0,
        \TransitionVar_3} & \land 
    \Connected(\AcaOrBc{}, \Filter) \\
    \land 
    \Connected(\SomethingCSomething{}, \Filter) \\
    \land 
    \Image{}_{\Tuple{\AcaOrBc{},\SomethingCSomething{}}, \Map}(\Filter, 
    \VectorLiteral{\TransitionVar_{7a} + \TransitionVar_{9a}, 0, \TransitionVar_{7c} + \TransitionVar_{9c} + 1}) & \land \TransitionVar_{7a} + \TransitionVar_{9a} > 0
  \end{array}}
  }
  \infer1[\EquationReasoning]{
    \begin{aligned}
      \VectorLiteral{a, b, c} = \VectorLiteral{
          2 - 2\TransitionVar_6,
          2\TransitionVar_4 + 3\TransitionVar_6,
          \TransitionVar_3 + \TransitionVar_6
        } \land
        \\
        \VectorLiteral{a, b, c} = \VectorLiteral{
            \TransitionVar_{7a} + \TransitionVar_{9a},
            \TransitionVar_{7b} + \TransitionVar_{9b},
            \TransitionVar_{7c} + \TransitionVar_{9c} + 1
          } \land \\
      \Connected(\AcaOrBc{}, \Filter) \land 
      \Connected(\SomethingCSomething{}, \Filter) \land \\
      \Image{}_{\Tuple{\AcaOrBc{},\SomethingCSomething{}}, \Map}(\Filter, 
      \VectorLiteral{a, b, c}) \land a > b
    \end{aligned}
  }
  \infer1[\ExpandM]{\Image{}_{\Tuple{A', B}, \Map}(\Filter, \VectorLiteral{l_A, i, n, l_B})}
\end{prooftree}
\caption{A derivation for \Calculus{} on the Parikh image of strings with more a's than b's in the product of $\AcaOrBc{}$ and $\SomethingCSomething{}$.}\label{fig:derivation:multi}
\end{figure}

>>>>

We reuse the transition variable associations of our previous
\cref{sec:single:example}, from \cref{eq:single-filter}, but add the definitions
for $B$ shown in \cref{eq:multi-filter}:
\begin{equation}\label{eq:multi-filter}
    \begin{aligned}
        \Filter(\FromLabelTo{A}{a / l_{B+}}{F_B} ) = \TransitionVar_6\\
        \Filter(\FromLabelTo{A}{c / l_{B+}}{A} ) = \TransitionVar_7\\
        \Filter(\FromLabelTo{B}{b / l_{B+}}{I_B})= \TransitionVar_8 \\
        \Filter(\FromLabelTo{B}{c / l_{B+}}{B})= \TransitionVar_9\\
        \Filter(\FromLabelTo{B}{c / l_{B+}}{F_B})= \TransitionVar_{10}\\
        \Filter(\FromLabelTo{I_B}{a / l_{B+}}{A})= \TransitionVar_{11}\\
        \Filter(\FromLabelTo{I_B}{b / l_{B+}}{B})= \TransitionVar_{12}\\
        \FlowEq(B, \Filter) =  \begin{aligned}
            1 + \TransitionVar_{12}   &= \TransitionVar_{11} + \TransitionVar_{12}\\
            \land \TransitionVar_{11} &= \TransitionVar_6\\
            \land \TransitionVar_{12} &= \TransitionVar_8 + \TransitionVar_{10}
        \end{aligned}
    \end{aligned}
\end{equation}

@@

\begin{figure}
  \centering
\begin{prooftree}
  % LEFT TREE  ->
  \hypo{l}% here x_12 is zero
  % RIGHT TREE ->
  \hypo{r} % here x_12 is one
  \infer2[\Split{} $\TransitionVar_{12} = 0$]{\text{FINAL STEP OF STUMP}}
  \infer1[\EquationReasoning: probably something good]{
    \begin{aligned}
      l_A = \sum_{k = 6}^{k = 12} \TransitionVar_k\\
      \land \TransitionVar_2 = l_A - 3\\
      \land 1 + \TransitionVar_{12}   = \TransitionVar_{11} + \TransitionVar_{12}\\
      \land \TransitionVar_{11} = \TransitionVar_6\\
      \land \TransitionVar_{12} = \TransitionVar_8 + \TransitionVar_{10}\\
      \land \Connected(B, \Filter) \\
      \land \Image{}_{\Tuple{A',B}, \Map}(\Filter, \VectorLiteral{l_A, 1, 2, l_A}) 
    \end{aligned}
  }
  \infer1[\EquationReasoning: expand $\FlowEq$]{
    \begin{aligned}
      l_A = \sum_{k = 6}^{k = 12} \TransitionVar_k\\
      \land \TransitionVar_2 = l_A - 3\\
      \land \FlowEq(B, \Filter) \\
      \land \Connected(B, \Filter) \\
      \land \Image{}_{\Tuple{A',B}, \Map}(\Filter, \VectorLiteral{l_A, 1, 2, l_A}) 
    \end{aligned}
  }
  \infer1[Repeat \cref{fig:derivation:single}, $l_A = l_B$.]{
    \begin{aligned}
      l_B = \sum_{k = 6}^{k = 12} \TransitionVar_k\\
      \land i = \TransitionVar_5\\
      \land l_A = \sum_{k = 1}^{k = 5} \TransitionVar_k\\
      \land n = \TransitionVar_4 + \TransitionVar_1\\  
      \land \FlowEq(B, \Filter) \\
      \land \Connected(A', \Filter) \\
      \land \Connected(B, \Filter) \\
      \land \Image{}_{\Tuple{A',B}, \Map}(\Filter, \VectorLiteral{l_A, i, n, l_B}) 
      \land l_A = l_B\\
      \land l_A - i - n > 0\\
      \land 0 < i < n
    \end{aligned}
  }
  \infer1[\ExpandM]{
    \Image{}_{\Tuple{A', B}, \Map}(\Filter, \VectorLiteral{l_A, i, n, l_B})
    \land l_A = l_B % Implied constraint
    \land l_A - i - n > 0
    \land 0 < i < n
    }
\end{prooftree}
\caption{A derivation for \Calculus{} on the Parikh image strings for the
constraints of \cref{ex:string-constraints}. Note the constant propagation
for~$n$!}\label{fig:derivation:multi}
\end{figure}
